MT @HanBaldwin: Fightin OOVs in German \#twitter

### 2013-08-05 darmstadt_gscl_2013/article_twitter_normalization.tex
small updates in article

  This article gives an overview of existing approaches to the problem
  of out-of-vocabulary (OOV) tokens and noisiness phenomena in natural
  language texts. These approaches are classified with regard to the
  size of text spans and knowledge inference mechanisms which they
  rely on in their work. Additionally to that, we conduct quantitative
  and qualitative analyses of unknown words in German Twitter
  messages, in order to see how relevant the OOV and text
  normalization problems are for this particular kind of micro-texts
  and what the characteristics of these OOVs are. In a concluding
  step, we present a set of ad-hoc techniques which are supposed to
  tackle some of the most prominent disturbing effects found during
  these analyses and show how this set of techniques helps us lower
  the average rate of out-of-vocabulary tokens in Twitter messages and
  how this lower OOV-rate in turn helps improve the quality of
  automatic part-of-speech tagging.

@@

In the next Section we will give a short overview of recent scientific
approaches to the problem of tackling out-of-vocabulary words in
non-standard texts. After that, in Section \ref{error:analysis} we
will analyze which types of noisiness phenomena are especially
characteristic for German Twitter. Section \ref{normalization} will
subsequently describe an automatic procedure for mitigating some of
the most prominent of those effects. In a concluding step, we will
perform an evaluation of the results of this procedure and give
further suggestions for future research.

@@

Before proceeding with the description of existing methods for noisy
text normalization (NTN), we first would like to define the criteria
by which these methods could be classified. It should be noted that
there already exist several classifications of NTN approaches
including Kukich (\cite{kukich}), Kobus et al. (\cite{kobus}), and
Sproat et al. (\cite{sproat}).
%% But all of these classifications seem to share one significant common
%% shortcoming, namely they try to involve several independent criteria
%% within one classification scheme.

Kukich (\cite{kukich}), for example, divided NTN techniques into six
classes:
\begin{enumerate}
\item minimum edit distance techniques;
\item similarity key techniques;
\item rule-based techniques;
\item \textit{n}-gram based techniques;
\item probabilistic techniques;
\item neural nets.
\end{enumerate}
Kobus (\cite{kobus}), on the contrary, referred to NTN methods as
\textit{metaphors} and split them into the following three groups:
\begin{enumerate}
\item ``spell checking'' metaphor;
\item ``translation'' metaphor;
\item ``speech recognition'' metaphor.
\end{enumerate}

@@

With these two classifications, we will now try to present and group
together approaches to the NTN task which appeared in the literature
in recent years.

The earlier works on NTN commonly relied on either purely graphematic
or phonographematic levels of segmentation for deriving normalization
variants of incorrectly spelled words. To purely graphematic systems
belong methods suggested by Brill and Moore (\cite{brill}), Sproat et
al. (\cite{sproat}), and Clark (\cite{alexander-clark}). As
phonographematic approaches one could regard works done by Toutanova
and Moore (\cite{toutanova}), Choudhury et al. (\cite{choudhury}),
Cook and Stevenson (\cite{cook}) etc. With regard to the type of
information inference, practically all of these methods were
supervised with the exception of Cook and Stevenson (\cite{cook}) who
claimed to use an unsupervised technique.

@@

Here once again the question of classification arose. This time with
regard to the reasons, why different tokens could have been omitted
from corresponding applications' dictionaries. In this respect,
division into following groups seemed to us to be appropriate:
\begin{enumerate}
  \item \textbf{Objective limitedness of machine-readable dictionary
    (MRD)}. Among this group, we counted words of basic vocabulary
    which did not get into applications' MRD either because they
    supposedly were rare or because they did not exist at the time
    when dictionaries were created. Another reason for the inclusion
    in this type was the belonging of a word to an open lexical or
    part-of-speech class (like, for example, named entities or
    interjections) which are often omitted from MRDs due to
    the impossibility to fully cover these classes;\label{dict}
  \item \textbf{Sloppiness of users' input}. In the scope of this
    second group, we considered incorrect spellings of words
    encountered in text;\label{spell}
  \item \textbf{Stylistic specifics of text genre}. This group
    comprised words which could be considered as illegal from the
    point of view of standard language but were perfectly valid terms
    in the domain of web discourse or more specifically in Twitter
    communication.\label{style}
\end{enumerate}

@@

We subdivided the class \ref{dict} into the following subclasses:
\begin{enumerate}
\item regular German words, e.g. \textit{aufm},
  \textit{losziehen};\label{regular}
\item compounds, e.g.  \textit{Altwein},
  \textit{Amtsapothekerin};\label{compound}
\item abbreviations, e.g. \textit{NBG}, \textit{OL};\label{abbr}
\item interjections, e.g.  \textit{aja}, \textit{haha};\label{inj}
\item named entities, with subclasses:\label{ne}
  \begin{enumerate}
  \item persons, e.g.  \textit{Ahmadinedschad}, \textit{Schweiger};
  \item geographic locations, e.g.  \textit{Biel}, \textit{Limmat};
  \item companies, e.g. \textit{Apple}, \textit{Facebook};
  \item product names, e.g. \textit{iPhone}, \textit{MacBook};
  \end{enumerate}
\item neologisms, with subclasses:\label{neolog}
  \begin{enumerate}
    \item newly coined German terms, e.g. \textit{entfolgen},
      \textit{gegoogelt};\label{new}
    \item loanwords, e.g. \textit{Community},
      \textit{Stream};\label{loan}
  \end{enumerate}
\item and, finally, foreign words like \textit{is} or \textit{now}
  which in contrast to \ref{loan} were not mentioned in any existing
  German lexica and did not comply with inflectional rules of German
  grammar.\label{fw}
\end{enumerate}
Though this division is admittedly arbitrarily to a certain degree and
also has the disadvantage of simultaneously involving different
linguistic criteria, the underlying notion here was simple -- valid
words could have been omitted from an MRD either due to the
limitations of developers' capacities (group \ref{regular}), active
word formation processes or lexical productivity of the language
itself (groups \ref{compound} through \ref{new}) or also due to
language's openness to foreign language systems (groups \ref{loan} and
\ref{fw}).

In Table \ref{table:mrd}, percentage figures for each of the above
subgroups are shown. We have considered OOV-distributions for both
\texttt{hunspell} and \texttt{TreeTagger}. For each of them, we
estimated the percentage of a particular subclass with regard to the
total number of occurrences of OOV-tokens (column ``\totaloov{}'') as
well as with regard to their percentage rate in the list of only
unique OOVs disregarding their frequencies (column ``\uniqoov{}'').
\begin{table}
  \caption{Distribution of OOV words belonging to the class
    ``Objective limitedness of MRD''\label{table:mrd}}
  \begin{tabular}{p{0.4\textwidth}*{4}{>{\centering\arraybackslash}p{0.15\textwidth}}}
    \hline\noalign{\smallskip}
    \multirow{2}{*}{OOV subclass} & %
    \multicolumn{2}{c}{\texttt{hunspell}} & %
    \multicolumn{2}{c}{\texttt{TreeTagger}}\\
    & \totaloov{} & \uniqoov{} & \totaloov{} & \uniqoov{}\\
    \noalign{\smallskip} \hline
    regular German words & 7.82 & 8.78 & 2.8 & 3.49\\
    compounds & 1.21 & 2.42 & 2.51 & 4.55\\
    abbreviations & 3.98 & 4.77 & 3.27 & 3.44\\
    interjections & 5.95 & 4.54 & 5.58 & 4.29\\
    person names & 4.73 & 6.41 & 2.32 & 3.47\\
    geographic locations & 1.5 & 2.53 & 1.16 & 1.88\\
    company names & 2.27 & 2.84 & 4.35 & 3.01\\
    product names & 2.13 & 2.57 & 2.45 & 3.23\\
    newly coined terms & 1.35 & 1.31 & 3.33 & 2.38\\
    loanwords & 3.68 & 4.03 & 3.29 & 2.87\\
    foreign words & 11.5 & 13.76 &  9.57 & 10.91\\\hline
    {\bfseries total} & 46.12 & 53.96 & 40.63 & 43.52\\
    \noalign{\smallskip} \hline
  \end{tabular}
\end{table}

Similarly to the class \ref{dict}, we subdivided the group
``Sloppiness of users' input'' into the following subgroups:
\begin{enumerate}
  \item insertions, e.g. \textit{dennen} instead of \textit{denen};
  \item deletions, e.g. \textit{scho} instead of \textit{schon};
  \item substitutions, e.g. \textit{fur} instead of \textit{f\"ur};
\end{enumerate}
according to the type of operation which led to a particular spelling
mistake. In cases when multiple different operations were involved
simultaneously, we explicitly marked each of these operations in our
data. Statistical distribution of these subclasses is shown in Table
\ref{table:spell}.
\begin{table}
  \caption{Distribution of OOV words belonging to the class
    ``Sloppiness of users' input''\label{table:spell}}
  \begin{tabular}{p{0.4\textwidth}*{4}{>{\centering\arraybackslash}p{0.15\textwidth}}}
    \hline\noalign{\smallskip}
    \multirow{2}{*}{OOV subclass} & %
    \multicolumn{2}{c}{\texttt{hunspell}} & %
    \multicolumn{2}{c}{\texttt{TreeTagger}}\\
    & \totaloov{} & \uniqoov{} & \totaloov{} & \uniqoov{}\\
    \noalign{\smallskip} \hline
    insertions & 0.49 & 1 & 0.18 & 0.34 \\
    deletions & 8.44 & 6.38 & 6.52 & 5.27 \\
    substitutions & 2.17 & 3.37 & 1.11 & 1.2 \\\hline
    {\bfseries total} & 11.1 & 10.75 & 7.81 & 6.81\\
    \noalign{\smallskip} \hline
  \end{tabular}
\end{table}

As is clear from the table, deletions are by far the most common type
of misspellings. The reasons for that are on the one hand either
relatively frequent omissions of characters made by user or even more
often automatic truncations of too long messages which were performed
by Twitter service itself.\footnote{As is generally known, Twitter
  imposes a strong restriction on the length of posted messages which
  can be no longer than 140 characters. Upon exceeding this length,
  tweets get automatically truncated to the maximal allowed length.}

Finally, the third group -- ``Stylistic specifics of text genre'' --
was subdivided into the subgroups:
\begin{enumerate}
  \item hashtags, e.g. \textit{\#Kleinanzeigen};
  \item @-tokens, e.g. \textit{@ZDFonline};
  \item links, e.g. \textit{http://t.co};
  \item smileys, e.g. \textit{:-P};
  \item slang, e.g. \textit{OMG}.
\end{enumerate}
according to the lexical or syntactic class which these tokens
belonged to. Apart from lexical expressions pertaining to the genre of
Internet-based communication, we also marked as \textit{slang}
spelling variants of standard language words which could be regarded
as their corresponding colloquial equivalents. As such equivalents, we
considred spellings like \textit{ned} instead of \textit{nicht} or
\textit{grad} instead of \textit{gerade}. These cases were marked by
us both as misspellings and as slang. A detailed statistics on the
above subgroups of class 3 is shown in Table \ref{table:style}:
\begin{table}
  \caption{Distribution of OOV words belonging to the class
    ``Stylistic specifics of text genre''\label{table:style}}
  \begin{tabular}{p{0.4\textwidth}*{4}{>{\centering\arraybackslash}p{0.15\textwidth}}}
    \hline\noalign{\smallskip}
    \multirow{2}{*}{OOV subclass} & %
    \multicolumn{2}{c}{\texttt{hunspell}} & %
    \multicolumn{2}{c}{\texttt{TreeTagger}}\\
    & \totaloov{} & \uniqoov{} & \totaloov{} & \uniqoov{}\\
    \noalign{\smallskip} \hline
    hashtags & 7.35 & 6.18 & 13.06 & 10.59\\
    @-tokens & 13.02 & 20.23 & 16.19 & 21.91\\
    links & 2.43 & 0.4 & 4.89 & 6.07\\
    smileys & 2 & 0.73 & 6.88 & 1.2\\
    slang & 16.22 & 5.27 & 6.94 & 4.77\\\hline
    {\bfseries total} & 41.02 & 32.81 & 47.96 & 44.54\\
    \noalign{\smallskip} \hline
  \end{tabular}
\end{table}

>>>>

  This article gives an overview of existing approaches to the problem
  of out-of-vocabulary (OOV) tokens and noisiness phenomena in natural
  language texts. These approaches are classified with regard to the
  size of text spans and knowledge inference mechanisms which they
  rely on in their work. Additionally to that, we conduct quantitative
  and qualitative analyses of unknown words in German Twitter
  messages, in order to see how relevant the OOV and text
  normalization problems are for this particular kind of micro-texts
  and what the characteristics of these OOVs are. In a concluding
  step, we present a set of ad-hoc techniques which are supposed to
  tackle some of the most prominent disturbing effects found during
  the analyses and show how this set of techniques helps us lower the
  average rate of out-of-vocabulary tokens in Twitter messages and how
  this lower OOV-rate in turn helps improve the quality of automatic
  part-of-speech tagging.

@@

In the next Section we will give a short overview of existing
scientific approaches to the problem of tackling out-of-vocabulary
words in non-standard texts. After that, in Section
\ref{error:analysis} we will analyze which types of noisiness
phenomena are especially characteristic for German Twitter. Section
\ref{normalization} will subsequently describe an automatic procedure
for mitigating some of the most prominent of those effects. In a
concluding step, we will perform an evaluation of the results of this
procedure and give further suggestions for future research.

@@

Before proceeding with the description of recent methods for noisy
text normalization (NTN), we first would like to define the criteria
by which these methods could be classified. It should be noted that
there already exist several classifications of NTN approaches
including, for example, Kukich (\cite{kukich}) and Kobus et
al. (\cite{kobus}).

So, Kukich (\cite{kukich}), for instance, divided all NTN techniques
into six classes:
\begin{enumerate}
\item minimum edit distance techniques;
\item similarity key techniques;
\item rule-based techniques;
\item \textit{n}-gram based techniques;
\item probabilistic techniques;
\item neural nets.
\end{enumerate}
Kobus (\cite{kobus}), on the contrary, referred to NTN methods as
\textit{metaphors} and split them into the following three groups:
\begin{enumerate}
\item ``spell checking'' metaphor;
\item ``translation'' metaphor;
\item ``speech recognition'' metaphor.
\end{enumerate}

@@

Based on these two classifications, recently proposed approaches to
the NTN task could be grouped together as follows.

The earlier works on NTN commonly relied on either purely graphematic
or phonographematic levels of segmentation for deriving normalization
variants of incorrectly spelled words. To purely graphematic systems
belong methods suggested by Brill and Moore (\cite{brill}), Sproat et
al. (\cite{sproat}), and Clark (\cite{alexander-clark}). As
phonographematic approaches one could regard works done by Toutanova
and Moore (\cite{toutanova}), Choudhury et al. (\cite{choudhury}),
Cook and Stevenson (\cite{cook}) etc. With regard to the type of
information inference, practically all of these methods were
supervised with the exceptions of Cook and Stevenson (\cite{cook}) who
claimed to use an unsupervised technique, and Sproat et
al. (\cite{sproat}) who described both a supervised and an
unsupervised technique.

@@

Here once again the question of classification arose. This time with
regard to the reasons, why different tokens could have been omitted
from corresponding applications' dictionaries. In this respect,
division into following groups seemed to us to be appropriate:
\begin{enumerate}
  \item \textbf{Objective limitedness of machine-readable dictionary
    (MRD)}. Among this group, we counted words of basic vocabulary
    which did not get into applications' MRD either because they
    supposedly were rare or because they did not exist at the time
    when dictionaries were created. Another reason for inclusion in
    this type was the belonging of a word to an open lexical or
    part-of-speech class (like, for example, named entities or
    interjections) which are often omitted from MRDs due to the
    impossibility to fully cover them;\label{dict}
  \item \textbf{Sloppiness of users' input}. In the scope of this
    second group, we considered incorrect spellings of words
    encountered in text;\label{spell}
  \item \textbf{Stylistic specifics of text genre}. This group
    comprised words which could be considered as illegal from the
    point of view of standard language but were perfectly valid terms
    in the domain of web discourse or more specifically in Twitter
    communication.\label{style}
\end{enumerate}

@@

We subdivided class \ref{dict} into the following subclasses:
\begin{enumerate}
\item regular German words, e.g. \textit{aufm},
  \textit{losziehen};\label{regular}
\item compounds, e.g.  \textit{Altwein},
  \textit{Amtsapothekerin};\label{compound}
\item abbreviations, e.g. \textit{NBG}, \textit{OL};\label{abbr}
\item interjections, e.g.  \textit{aja}, \textit{haha};\label{inj}
\item named entities, with subclasses:\label{ne}
  \begin{enumerate}
  \item persons, e.g.  \textit{Ahmadinedschad}, \textit{Schweiger};
  \item geographic locations, e.g.  \textit{Biel}, \textit{Limmat};
  \item companies, e.g. \textit{Apple}, \textit{Facebook};
  \item product names, e.g. \textit{iPhone}, \textit{MacBook};
  \end{enumerate}
\item neologisms, with subclasses:\label{neolog}
  \begin{enumerate}
    \item newly coined German terms, e.g. \textit{entfolgen},
      \textit{gegoogelt};\label{new}
    \item loanwords, e.g. \textit{Community},
      \textit{Stream};\label{loan}
  \end{enumerate}
\item and, finally, foreign words like \textit{is} or \textit{now}
  which in contrast to \ref{loan} were not mentioned in any existing
  German lexica and did not comply with inflectional rules of German
  grammar.\label{fw}
\end{enumerate}
Though this division is admittedly arbitrarily to a certain degree and
also has the disadvantage of simultaneously involving different
linguistic criteria, the underlying notion here was simple -- valid
words could have been omitted from an MRD either due to the
limitations of developers' capacities (group \ref{regular}), active
word formation processes or lexical productivity of the language
itself (groups \ref{compound} through \ref{new}) or also due to
language's openness to foreign language systems (groups \ref{loan} and
\ref{fw}).

In Table \ref{table:mrd}, percentage figures for each of the above
subgroups are shown. We have considered OOV-distributions for both
\texttt{hunspell} and \texttt{TreeTagger}. For each of them, we
estimated the percentage of a particular subclass with regard to the
total number of occurrences of all OOV-tokens (column ``\totaloov{}'')
as well as with regard to their percentage rate in the list of only
unique OOVs disregarding the frequencies (column ``\uniqoov{}'').
\begin{table}
  \caption{Distribution of OOV words belonging to the class
    ``Objective limitedness of MRD''\label{table:mrd}}
  \begin{tabular}{p{0.4\textwidth}*{4}{>{\centering\arraybackslash}p{0.15\textwidth}}}
    \hline\noalign{\smallskip}
    \multirow{2}{*}{OOV subclass} & %
    \multicolumn{2}{c}{\texttt{hunspell}} & %
    \multicolumn{2}{c}{\texttt{TreeTagger}}\\
    & \totaloov{} & \uniqoov{} & \totaloov{} & \uniqoov{}\\
    \noalign{\smallskip} \hline
    regular German words & 7.82 & 8.78 & 2.8 & 3.49\\
    compounds & 1.21 & 2.42 & 2.51 & 4.55\\
    abbreviations & 3.98 & 4.77 & 3.27 & 3.44\\
    interjections & 5.95 & 4.54 & 5.58 & 4.29\\
    person names & 4.73 & 6.41 & 2.32 & 3.47\\
    geographic locations & 1.5 & 2.53 & 1.16 & 1.88\\
    company names & 2.27 & 2.84 & 4.35 & 3.01\\
    product names & 2.13 & 2.57 & 2.45 & 3.23\\
    newly coined terms & 1.35 & 1.31 & 3.33 & 2.38\\
    loanwords & 3.68 & 4.03 & 3.29 & 2.87\\
    foreign words & 11.5 & 13.76 &  9.57 & 10.91\\\hline
    {\bfseries total} & 46.12 & 53.96 & 40.63 & 43.52\\
    \noalign{\smallskip} \hline
  \end{tabular}
\end{table}

Similarly to class \ref{dict}, we subdivided the group ``Sloppiness of
users' input'' into the following subgroups:
\begin{enumerate}
  \item insertions, e.g. \textit{dennen} instead of \textit{denen};
  \item deletions, e.g. \textit{scho} instead of \textit{schon};
  \item substitutions, e.g. \textit{fur} instead of \textit{f\"ur};
\end{enumerate}
according to the type of operation which led to a particular spelling
mistake. In cases when multiple different operations were involved
simultaneously, we explicitly marked each of these operations in our
data. Statistical distribution of these subclasses is shown in Table
\ref{table:spell}.
\begin{table}
  \caption{Distribution of OOV words belonging to the class
    ``Sloppiness of users' input''\label{table:spell}}
  \begin{tabular}{p{0.4\textwidth}*{4}{>{\centering\arraybackslash}p{0.15\textwidth}}}
    \hline\noalign{\smallskip}
    \multirow{2}{*}{OOV subclass} & %
    \multicolumn{2}{c}{\texttt{hunspell}} & %
    \multicolumn{2}{c}{\texttt{TreeTagger}}\\
    & \totaloov{} & \uniqoov{} & \totaloov{} & \uniqoov{}\\
    \noalign{\smallskip} \hline
    insertions & 0.49 & 1 & 0.18 & 0.34 \\
    deletions & 8.44 & 6.38 & 6.52 & 5.27 \\
    substitutions & 2.17 & 3.37 & 1.11 & 1.2 \\\hline
    {\bfseries total} & 11.1 & 10.75 & 7.81 & 6.81\\
    \noalign{\smallskip} \hline
  \end{tabular}
\end{table}

As is clear from the table, deletions are by far the most common type
of misspellings. The reasons for that are on the one hand either
relatively frequent omissions of characters made by users or even more
often automatic truncation of too long messages which is performed by
Twitter service itself.\footnote{As is generally known, Twitter
  imposes a strong restriction on the length of posted messages which
  can be no longer than 140 characters. Upon exceeding this length,
  tweets get automatically truncated to the maximal allowed length.}

Finally, the third group -- ``Stylistic specifics of text genre'' --
was subdivided into the subgroups:
\begin{enumerate}
  \item hashtags, e.g. \textit{\#Kleinanzeigen};
  \item @-tokens, e.g. \textit{@ZDFonline};
  \item links, e.g. \textit{http://t.co};
  \item smileys, e.g. \textit{:-P};
  \item slang, e.g. \textit{OMG}.
\end{enumerate}
according to the formal or lexical class which these tokens belonged
to. Additionally, we also considered as \textit{slang} spelling
variants of standard language words which could be regarded as their
corresponding colloquial equivalents like, for example, \textit{ned}
instead of \textit{nicht} or \textit{grad} instead of \textit{gerade}.

These cases were marked by
us both as misspellings and as slang. A detailed statistics on the
above subgroups of class 3 is shown in Table \ref{table:style}:
\begin{table}
  \caption{Distribution of OOV words belonging to the class
    ``Stylistic specifics of text genre''\label{table:style}}
  \begin{tabular}{p{0.4\textwidth}*{4}{>{\centering\arraybackslash}p{0.15\textwidth}}}
    \hline\noalign{\smallskip}
    \multirow{2}{*}{OOV subclass} & %
    \multicolumn{2}{c}{\texttt{hunspell}} & %
    \multicolumn{2}{c}{\texttt{TreeTagger}}\\
    & \totaloov{} & \uniqoov{} & \totaloov{} & \uniqoov{}\\
    \noalign{\smallskip} \hline
    hashtags & 7.35 & 6.18 & 13.06 & 10.59\\
    @-tokens & 13.02 & 20.23 & 16.19 & 21.91\\
    links & 2.43 & 0.4 & 4.89 & 6.07\\
    smileys & 2 & 0.73 & 6.88 & 1.2\\
    slang & 16.22 & 5.27 & 6.94 & 4.77\\\hline
    {\bfseries total} & 41.02 & 32.81 & 47.96 & 44.54\\
    \noalign{\smallskip} \hline
  \end{tabular}
\end{table}
